A Cross-Domain Benchmark for Active Learning

### 2024-06-04 doc/al_benchmark_paper.tex
Notes on Orcale

\section{A Greedy Oracle Algorithm}\label{sec:oracle}
An oracle method for AL tries to find the oracle set $\mathcal{O}_b$ for a given dataset, model, and training procedure that induces the highest AUC score for a given budget.
However, due to the combinatorial nature of the problem, it is computationally infeasible for realistic datasets.
Hence, previous works have proposed approximations to this oracle sequence. 
\cite{zhou2021towards} used simulated annealing to search for the optimal subset and used the best  solution found after a fixed time budget. 
Even though their reported performance curves display a significant lift over all other acquisition functions, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs).
In this paper, we propose a greedy oracle algorithm that constructs an approximation of the optimal set in an iterative fashion.
Our oracle algorithm evaluates every data point $u_k = \operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$ in a subsample of unlabeled points by recovering the label, fitting the classifier $\hat y$ on $\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test performance.
The data point that incurs the largest lift in test performance is selected and added to the labeled pool for that iteration.

@@

Due to the algorithms greedy nature (just considering the next point to pick), our oracle frequently encounters situations where every point in $u_k$ would incur a negative lift (worsening the test performance).
This can happen, for example, if the oracle picked a labeled set that enables the classifier to correctly classify a big portion of easy samples in the test set, but now fails to find the next \textbf{single} unlabeled point that would enable the classifier to succeed on one of the hard samples.
This leads to a situation, where no point can immediately incur an increase in test performance and therefore the selected data point can be considered random.
To circumvent this problem, we use our best-performing acquisition function (margin sampling \cite{wang2014new}) as a fallback option for the oracle.
Whenever the oracle does not find an unlabeled point that results in an increase in performance, it defaults to margin sampling in that iteration.
The resulting greedy algorithm constructs an approximation of the optimal labeled set that consistently outperforms all other algorithms by a significant margin, while requiring relatively low computational cost ($\mathcal{O}(B\tau)$).

@@

We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the best performing acquisition function per dataset (which is significant for AL settings) and we expect diminishing returns for larger $\tau$.
The pseudocode for our oracle can be found in App. \ref{app:pseudocode}.

@@

Even though our proposed algorithm is more efficient than other approaches, the computational costs for high budget datasets like Cifar10 and FashionMnist meant that we could not compute the oracle for all 10000 datapoints.
To still provide an oracle for these two datasets, we select two points per iteration instead of one and stop the oracle computation at a budget of 2000.
The rest of the curve is forecast with a simple linear regression that asymptotically approaches the upper bound performance of the dataset. 
A detailed description can be found in App. \ref{app:oracle_forecasting}.

>>>>

\section{A Greedy Oracle Algorithm}\label{sec:oracle} An oracle method for AL
tries to find the oracle set $\mathcal{O}_b$ for a given dataset,\todo{Versuchen
die Beste zu finden tut doch jede AL-Methode! Vlt sowas wie: Ist ein Verfahren,
dass zwar die Optimale Lösung findet, aber dabei keine Rücksicht auf Laufzeiten nimmt} model, and
training procedure that induces the probably highest AUC score for a given budget.
However, due to the combinatorial nature of the problem, it is computationally
infeasible for realistic datasets. Hence, previous works have proposed
approximations to this oracle sequence. \cite{zhou2021towards} used simulated
annealing to search for the optimal subset and used the best  solution found
after a fixed time budget. \todo{Okay und warum ist das dann keine Aquisition funnktion}Even though their reported performance curves display
a significant lift over all other acquisition functions, we found the
computational cost of reproducing this oracle for all our datasets to be
prohibitive (The authors reported the search to take several days per dataset on
8 V100 GPUs). In this paper, we propose a greedy oracle algorithm that
constructs an approximation of the optimal set in an iterative fashion. Our
oracle algorithm evaluates every data point $u_k =
\operatorname{unif(\mathcal{U}) \quad k \in [1 \ldots \tau]}$\todo[inline]{Die Notation
stimmt so nicht, willst du sagen, dass du hier ein Subset der größe $\tau$
gesamplet hast uniform gesamplet hast? Dann schreibe das: "Our oracle algorithm
uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of the not already
labeled data points $\mathcal{U}^{(i)}$.
Then it recovers the label $y$ for each of the sampled $u \in \mathcal{U}_S$ and selects the
point $u \in \mathcal{U}_S$ for which the classifier $\hat{y}$ trained on
$\mathcal{L}^{(i)} \cup \{u\}$ has the maximal performance. } in a subsample of
unlabeled points by recovering the label, fitting the classifier $\hat y$ on
$\mathcal{L}^{(i)} \cup \{u_k\}$ and directly measuring the resulting test
performance. \todo{Okay, hier verstehe ich was ein Orcale ist. Eine Aquisition
funktion die schon das wissen über die Labels beim aquirieren nutzen kann?
Stimmt das?}The data point that incurs the largest lift in test performance is
selected and added to the labeled pool for that iteration.

@@

Due to the algorithms greedy nature (just considering the next point to pick),
our oracle frequently encounters situations where every point in $u_k$ would
incur a negative lift (worsening the test performance). This can happen, for
example, if the oracle picked a labeled set that enables the classifier to
correctly classify a big portion of easy samples in the test set, but now fails
to find the next \textbf{single} unlabeled point that would enable the
classifier to succeed on one of the hard samples. This leads to a situation,
where no point can immediately incur an increase in test performance and
therefore the selected data point can be considered random. To circumvent this
problem, we use our best-performing acquisition function (margin sampling
\cite{wang2014new}) as a fallback option for the oracle. Whenever the oracle
does not find an unlabeled point that results in an increase in performance, it
defaults to margin sampling in that iteration. The resulting greedy algorithm
constructs an approximation of the optimal labeled set that consistently
outperforms all other algorithms by a significant margin, while requiring
relatively low computational cost ($\mathcal{O}(B\tau)$).

@@

We fix $\tau = 20$ in this work, as this gave us an average lift of 5\% over the
best performing acquisition function per dataset (which is significant for AL
settings) and we expect diminishing returns for larger $\tau$. The pseudocode
for our oracle can be found in App. \ref{app:pseudocode}.

@@

Even though our proposed algorithm is more efficient than other approaches, the
computational costs for high budget datasets like Cifar10 and FashionMnist meant
that we could not compute the oracle for all 10000 datapoints. To still provide
an oracle for these two datasets, we select two points per iteration instead of
one and stop the oracle computation at a budget of 2000. The rest of the curve
is forecast with a simple linear regression that asymptotically approaches the
upper bound performance of the dataset. A detailed description can be found in
App. \ref{app:oracle_forecasting}.\todo[inline]{Approached heißt konvergiert,,
oder? Dann muss das doch nicht stimmen. Ich würde nach "linear regression" einen
Punkt machen und den Teil danach löschen.}
